---
title: "ClassAssignment8"
author: "Ben Pelczynski"
date: "2024-07-25"
output: html_document
---

```{r}
library(tidyverse)
library(ggcorrplot)
library(car)
credit_data <- read.csv("credit_data.csv")
```


```{r}
cor_mat <- cor(credit_data[2:7])
ggcorrplot(cor_mat, lab=TRUE, type="lower")
```

Based on this matrix plot, there are issues of multicollinearity with Limit and Rating specifically having a 1-to-1 correlation (much greater than the threshold of 0.8).
```{r}
model <- lm(Balance~Income+Limit+Rating+Cards+Age+Education, data = credit_data)
vif(model)
```

Based on this VIF statement, we will rebuild the model without Rating, the variable with the highest VIF.
```{r}
model <- lm(Balance~Income+Limit+Cards+Age+Education, data = credit_data)
vif(model)
```

All VIFs fall into acceptable parameters in the the rebuilt model.

```{r}
new_data <- data.frame(Income = c(175, 48, 100, 134, 89), 
                       Limit = c(9000, 2500, 5500, 10000, 7800), 
                       Cards = c(2, 4, 7, 1, 6), 
                       Age = c(33, 44, 68, 50, 100), 
                       Education = c(18, 13, 15, 17, 19))
predict(model, newdata = new_data, interval = "prediction", level = .95)
```

We can say with a 95% level of confidence that a Balance, given the reasonable values I inserted, would fall between the lwr and upr limits given by the prediction model. For example, a 68-year old with 7 cards, a credit limit of 5500, an income of \$100k, and an education level of 15, would have a balance between \$63.21 and \$715.61.
```{r}
predict(model, newdata = new_data, interval = "confidence", level = .95)

```

We can say with a 95% level of confidence that the mean Balance of population, given the reasonable values I inserted, would fall between the lwr and upr limits given by the prediction model. For example, the population of a 68-year olds with 7 cards, a credit limit of 5500, an income of \$100k, and an education level of 15, would have an average balance between \$326.64 and \$452.18.